---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# twfeBenchmark

<!-- badges: start -->
<!-- badges: end -->

The goal of twfeBenchmark is to provide a Shiny application as an exploration tool for understanding two-way fixed effect estimations.
By letting people experiment with small designs, we hope that they will have a better understanding of the recent results in the literature.
Indeed, although these types of models were widely used in the literature as a causal analysis tool, the basic ones have multiple limitations that could yield to misleading results.

## Install and Run the app

You can clone the project, and run in the R console:

```{r example, eval = FALSE}
# Option 1:
#   Use this command if you don't want to install the project as a package.
#   This should load all the functions of the project.
devtools::load_all()

# Option 2
#   Use this command if you want to install the project as a package.
#   Only some of the functions of the project will be exposed.
devtools::install()

# This will launch the Shiny App
twfeBenchmark::run_twfe_app()
```

This should launch a Shiny application and open an interactive window.

## Learning more

More details about the theory and the content of the app can be seen in the vignettes.
