---
title: "Theory behind Two-way Fixed Effects Regression Models"
output: rmarkdown::html_vignette
bibliography: bibliography.bib
link-citations: true
vignette: >
  %\VignetteIndexEntry{Theory behind Two-way Fixed Effects Regression Models}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo = FALSE}
library(twfeBenchmark)
```

## Global overview of the recent literature


Difference in Difference is one of the most common strategy for estimating causal effects.
Linear regression with period and group fixed effects were widely used in the literature as an attempt to generalize this basic estimator for situations with multiple treatments, dynamic effects, etc...
Recent works have tried to analyse the theoretical properties of these models.
We recommend that you look at the survey done by @chaisemartin_haultfoeuille_twfe_survey_2023 if you want a global overview of what has been done.
This vignette will be focused on the aspects of the theory that could be experimented in the application.

## Goodman-Bacon decomposition

The original DiD is computed using two groups and two periods ($2 \times 2$).
To compute it, we simply have to compare the change in outcomes after the treatment of a group and compare it against the change in outcomes of a control group that was untreated at both period.
It can also be estimated by the following regression, using a treatment group dummy $T_i$ and a post-treatment period dummy $P_t$ like so:
$$
Y_{it} = \mu + \alpha T_i + \gamma P_t + \beta^{2 \times 2}T_i P_t + \varepsilon_{it}
$$

A natural idea leading to two-way fixed effects, is that we would obtain the same DiD estimator if we used the following regression:
$$
Y_{it} = \alpha_i + \gamma_t + \beta^{DD}D_{it} + \varepsilon_{it}
$$
however, it is now possible to use this regression for groups being treated at different time, or even for non-discrete treatments.
Can we interpret this $\beta^{DD}$ similarly to $\beta^{2 \times 2}$ ?

@goodman-bacon_did_2021 is one of the first in attempting to give a better understanding of what this basic two-way fixed effect model estimates^[we stay in the situation where the treatment is binary and staggered].
His main result is that we can interpret $\hat{\beta}^{DD}$ as a weighted sum of basic DiD estimators.
More precisely, denote by $n_k$ the proportion of individuals in the group $k$, and $\overline{D}_k$ the proportion of time this group spends treated in the data.
$$
\hat{\beta}^{DD} = \sum_{k < l} s_{kl} \left[\mu_{kl} \hat{\beta}^{2 \times 2, k}_{kl} + (1 - \mu_{kl})\hat{\beta}_{kl}^{2 \times 2, l}\right]
$$
where we sorted groups by their time of treatment and

$$
\begin{align}
  s_{kl} &\propto n_k n_l (\overline{D}_k - \overline{D}_l) (1 - (\overline{D}_k - \overline{D}_l)), \quad \sum_{k < l} s_{kl} = 1\\
  \mu_{kl} &= \frac{1 - \overline{D}_k}{1 - (\overline{D}_k - \overline{D}_l)}
\end{align}
$$

We consider here that the group $u$ of untreated individuals is like a group of individuals that are treated at an infinite horizon, and thus $\overline{D}_u = 0$.
In particular, we obtain that $\mu_{ku} = 1$ and the term $\hat{\beta}_{ku}^{2 \times 2, u}$ is unimportant.

Here, $\hat{\beta}_{kl}^{2 \times 2, k}$ corresponds to a basic DiD between groups $k$ and $l$, using $k$ as the treated group and $l$ as the control group^[to do so, we therefore only keep the dates before his treatment].
This term can already be quite surprising because it means we don't always use $u$ as the control group.
However it kind of make sense because $k$ is treated first.
Thus at the time $k$ is treated, $l$ is among the not-yet treated.
If we assume that there are no anticipations that would change $l$ outcomes before his treatment, you could legitimately argue that $l$ can be considered as a control group.

The more surprising part is that there is also $\hat{\beta}_{kl}^{2 \times 2, l}$, using this time $l$ as the treated group and $k$ as the control group^[to do so, we therefore only keep the dates after his treatment].
One situation where this could lead to problems is where $k$'s treatment has dynamic effects, i.e. when the effect change over time after his treatment.
Even if the parallel trend assumption holded before $k$'s treatment, this dynamic effect would make the outcome path of $k$ unusable as a control group.
In fact, this DiD could give a negative sign even if every treatments had positive effects on every groups.
This is why these DiD can be described as "forbidden comparisons".

Now what does the weights tell us ?
$\mu_{kl}$ corresponds to the relative weights of the "allowed comparison" against the "forbidden comparison".
We can rewrite it as 
$$
\mu_{kl} = \frac{1 - \overline{D}_k}{(1 - \overline{D}_k) + \overline{D}_l}
$$
where $1-\overline{D}_k$ is the share of pre-treatment time of group $k$, and $\overline{D}_l$ is the share of treatment time of group $l$.
Thus, the later $k$ is treated and the later $l$ is treated, the higher is the weight of the "allowed comparison".

Now if we want to compare global weights, we can compute 
$$
\begin{align}
s_{kl}\mu_{kl} &\propto n_k n_l (1 - \overline{D}_k)(\overline{D}_k - \overline{D}_l)\\
s_{kl}(1 - \mu_{kl}) &\propto n_k n_l (\overline{D}_k - \overline{D}_l)\overline{D}_l
\end{align}
$$
and check that in both cases, the third and fourth terms correspond respectively to the share of pre-treatment time and post-treatment time of the considered DiD.

Therefore we can conclude that the importance of a DiD will depend on two aspects of groups and periods sizes: 

  - *Pure size*: the bigger the group sizes and the considered periods, the bigger the weight. 
  - *Balance*: if the compared groups are roughly of the same size, and if the considered pre-treatment period and post-treatment period are roughly of the same size, the DiD will have a bigger weight.
